Analysis of the inference of ratings and rankings on Higher Order Networks with complex topologies

### 2025-09-03 draft/draft.tex
progress

Consider a set of $N=n+1$ teams labeled by $\{0,1,...,n\}$ to be rated and ranked.
For this aim, assume that some set of pairwise comparison scores $\psi_{ij}=-\psi_{ji}$ is provided for each $ij$ of different labels $i$ and $j$.
Informally, the score $\psi_{ij}$ should be positive if the pairwise comparison indicates that team $i$ should be rated higher than team $j$ and it should be non-positive otherwise.
One way to obtain rating scores somehow consistent with the pairwise comparisons scores, consist in finding ratings $w_i$ for $i=0,1,..,n$ such that minimize the sum of squared differences
$$
\sum_{ij:i<j} |\psi_{ij}-(w_i-w_j)|^2.
$$

>>>>

\subsection{HodgeRank for inference of team's ratings and rankings}

%Consider a set of $N=n+1$ teams labeled by $\{0,1,...,n\}$ to be rated and ranked.
%For this aim, assume that some set of pairwise comparison scores $\psi_{ij}=-\psi_{ji}$ is provided for each pair $ij$ of different teams $i$ and $j$.
%Informally, the score $\psi_{ij}$ should be positive if the comparison indicates that team $i$ should preferably be rated higher than team $j$ and the opposite otherwise.
%
%One way to infer rating scores somehow consistent with the pairwise comparisons scores $\phi_{ij}$, consist in finding ratings $w_i$ for $i=0,1,..,n$ that minimize the sum of squared differences
%$$
%\sum_{ij:i<j} |\psi_{ij}-(w_i-w_j)|^2.
%$$
%This is a well defined mathematical problem that can be conveniently reformulated in the language of cochains over simplicial complexes, conceiving a method for the inference of ratings and rankings that is named  HodgeRank~\cite{jiang2011statistical,
%perotti2025analysis}.

Consider a set of $N=n+1$ teams labeled $\{0,1,\dots,n\}$ to be rated and ranked.
Suppose that for each distinct pair of teams $i$ and $j$, a pairwise comparison score $\psi_{ij}=-\psi_{ji}$ is given.
Informally, $\psi_{ij}>0$ suggests that team $i$ should be rated higher than team $j$, while $\psi_{ij}<0$ suggests the opposite.

A natural approach to inferring rating scores consistent with the pairwise comparisons is to assign ratings $w_i$ (for $i=0,1,\dots,n$) that minimize the sum of squared discrepancies
$$
\sum_{i<j} \bigl(\psi_{ij}-(w_i-w_j)\bigr)^2.
$$
This yields a well-posed optimization problem that admits an elegant reformulation in the language of cochains over simplicial complexes.
The resulting framework for the inference of ratings and rankings is known as {\em HodgeRank}~\cite{jiang2011statistical, perotti2025analysis}.

%To define HodgeRank, it is convenient to introduce a simplicial complex composed of a 0-simplex $i$ for each team, a 1-simplex $ij$ for each pairwise comparison $\psi_{ij}$ between teams and a 2-simplex $ijk$ for each existing triple $(ij,jk,ki)$ in the set of 1-simplices.
%Recall here that $ij$ and $ji$ are both labels of the same simplex, as well as any permutation of $ijk$ is also a label of the same 2-simplex.

To define HodgeRank, an abstract simplicial complex has to be defined.
For this, let the vertex set be $V=\{0,1,\dots,n\}$, representing the teams, and let the edge set $E$ consist of the unordered pairs $\{i,j\}$ for which a pairwise comparison score $\psi_{ij}$ is available.
Then, the associated abstract simplicial complex is
$$
K = K_0 \cup K_1 \cup K_2,
$$
where
\begin{eqnarray}
K_0 &=& \{\{i\} : i \in V\}, \notag \\  
K_1 &=& \{\{i,j\} : \{i,j\} \in E\}, \notag \\
K_2 &=& \{\{i,j,k\} : \{i,j\},\{j,k\},\{k,i\} \in E\} \notag
\end{eqnarray}
are the sets of 0, 1 and 2 simplices, respectively.
Thus, each team corresponds to a 0-simplex, each comparison to a 1-simplex, and each closed triple of comparisons to a 2-simplex.
By convention, $\{i,j\}=\{j,i\}$ denotes the same 1-simplex, and any permutation of $\{i,j,k\}$ denotes the same 2-simplex.
